\section{Installation instruction of  syfala v7 toolchain}
\label{annex}
\label{install}
The Syfala toolchain is a compilation toolchain of Faust programs on FPGA. This document explains how to install and run the toolchain v7  (version without petalinux), on a linux\footnote{tested on Ubuntu 18.04 and Ubuntu 20.04 and arch linux} machine. In practice, installing the Syfala tool-chain  means:
\begin{itemize}
\item Installing the Faust compiler, see section~\ref{faust-install} below.
\item Creating a Xilinx account and downloading/installing the 2020.2 version of the Xilinx {\tt Vivado} toolchain: {\tt vitis\_hls}, {\tt vivado} and {\tt vitis}. See section~\ref{vitis-install} below.
\item Installing Vivado board files for Digilent boards, see section~\ref{board-file-install}
\item Installing udev rules to use JTAG connection, see section~\ref{board-file-install}  
\item Cloning the Syfala directory and running a simple example as explained in Section~\ref{sec-syfala}.
\end{itemize}
Section~\ref{hard} explains the hardware configuration of the Zybo board for Syfala and Section~\ref{bug} list all the important bugs encountered when building Syfala. If you encounter a bug during the installation, please see Section~\ref{bug}.


{\bf Ubuntu dependencies:} Syfala dependencies on Linux Ubuntu are the following:\\
\texttt{sudo apt install libncurses5 libtinfo-dev g++-multilib gtk2.0}

{\bf Warning:} You need approximately 50GB of disk space to install the tool chain, and a good connection. The installation take several hours.
%If the installer prompts a choice for which version to install, select the {\bf WebPack Edition} 
           
%% {\bf Warning} all the tools of Vivado come with shell scripts that set up your {\tt \$PATH} to use them. It is quite dangerous to source them in the {\tt .bashrc} file because it provides older version of important utilities (such as {\tt cmake} for instance). We strongly advise you to use a fonction defined in your {\tt .bashrc} file such as the following:
%% ~\\

%% \begin{boxedminipage}{\textwidth}
%% \begin{verbatim}
%%   function use_vitis
%%   {
%%     source $myXilinxToolDirectory/Vivado/2020.2/settings64.sh
%%     source $myXilinxToolDirectory/Vitis_HLS/2020.2/settings64.sh
%%     source $myXilinxToolDirectory/Vitis/2020.2/settings64.sh
%%   }
%% \end{verbatim}
%% \end{boxedminipage}

\section{Installing Faust}
\label{faust-install}
It is recommanded to clone Faust from the github repository: \url{https://github.com/grame-cncm/faust}:
\begin{verbatim}
  git clone https://github.com/grame-cncm/faust faust
  cd faust
  make
  sudo make install
\end{verbatim}
If you are using an older version of Syfala, you might need to use older version of Faust (see {\tt version} files in Syfala directory). The procedure is to get the commit number of the version you need here: \url{https://github.com/grame-cncm/faust/releases}. For instance, if you use Syfala v5.4, it requires Faust version 2.31.1 (at least), it commit number is:  32a2e92c955c4e057d424ab69a84801740d37920, then execute:
\begin{verbatim}
cd faust 
git checkout  32a2e92c955c4e057d424ab69a84801740d37920
make 
sudo make install
\end{verbatim}

\section{Installing {\tt Vivado}, {\tt Vitis} and {\tt Vitis\_hls} }
\label{vitis-install}


\begin{itemize}
\item
  Open an account on https://www.xilinx.com/registration
\item
  The Xilinx download page
  (https://www.xilinx.com/support/download.html) and browse to the
  2020.2 version. The page contains links for downloading the
  ``Xilinx\_Unified\_2020.2\_1118\_1232\_Lin64.bin'' (It is available
  for both Linux and Windows but Syfala compiles only on Linux).

  \begin{itemize}
  \item
    Download the Linux installer
    \texttt{Xilinx\_Unified\_2020.2\_1118\_1232\_Lin64.bin}
  \end{itemize}
\item
  execute
  \texttt{chmod\ a+x\ Xilinx\_Unified\_2020.2\_1118\_1232\_Lin64.bin}
\item
  execute \texttt{./Xilinx\_Unified\_2020.2\_1118\_1232\_Lin64.bin}

  \begin{itemize}
  \item
    We suggest to use the ``Download Image (Install Separately)''
    option. It creates a directory with a xsetup file to execute that
    you can reuse in case of failure during the installation
  \end{itemize}
\item
  execute \texttt{./xsetup}

  \begin{itemize}
  \item
    Choose to install \textbf{Vitis} (it will still install
    \textbf{Vivado}, \textbf{Vitis}, and \textbf{Vitis HLS}).
  \item
    It will need 110GB of disk space: if you uncheck \emph{Versal ACAP} and \emph{Alveo acceleration
    platform}, it will use less space and still work.
  \item
    Agree with everything and choose a directory to install
    (e.g.~\textasciitilde/Xilinx)
  \item
    Install and wait for hours\ldots{}
  \end{itemize}
\item
  Setup a shell function allowing to use the tools when necessary (add
  this to your \texttt{\textasciitilde{}/.bashrc},
  \texttt{\textasciitilde{}/.zshrc} or whatever you're currently using,
  replacing \texttt{\$XILINX\_ROOT\_DIR} by the directory you chose to
  install all the tools)

  \begin{itemize}
  \item
\begin{verbatim}
  export XILINX_ROOT_DIR=$HOME/Xilinx
\end{verbatim}
  \end{itemize}
\end{itemize}

Then Install missing Vivado board files for Digilent boards and drivers for linux (explained in Section~\ref{board-file-install} below).

\knownbug{You HAVE to read sections~\ref{localSetting} (locale setting) and \ref{2k22patch-install} (vivado 2022 bug patch). If you do not, you might end up with unpredictible behaviour of Vivado.
}




\section{Installing Vivado Board Files and Linux drivers}
\label{board-file-install}

\subsection{Cable drivers (Linux only)}
\label{sec-udev}
\begin{itemize}
\item
  go to:\\
  \texttt{\$XILINX\_ROOT\_DIR/}\\
  \texttt{Vivado/2020.2/data/xicom/cable\_drivers/lin64/install\_script/install\_drivers}\\
  directory
\item
  run \texttt{./install\_drivers}
\item
  run \texttt{sudo\ cp\ 52-xilinx-digilent-usb.rules\ /etc/udev/rules.d}, this
  allows \textbf{JTAG} connection through \textbf{USB}.
\end{itemize}

\subsection{Vivado Board Files for Digilent Boards}
{\bf Important}: This step is needed to enable vivado to generate code for the Zybo Z10

\begin{itemize}
\item
  download:\\
\href{https://github.com/Digilent/vivado-boards/archive/master.zip?\_ga=2.76732885.1953828090.1655988025-1125947215.1655988024}{https://github.com/Digilent/vivado-boards/archive/master.zip}
\item
  Open the folder extracted from the archive and navigate to its
  \texttt{new/board\_files} folder. You will be copying all of this
  folder's subfolders
\item
  go to
  \texttt{\$XILINX\_ROOT\_DIR/Vivado/2020.2/data/boards/board\_files}
\item
  \textbf{Copy} all of the folders found in vivado-boards
  \texttt{new/board\_files} folder and \textbf{paste} them into this
  folder
\end{itemize}

%
\subsection{Installing the 2022
patch}\label{2k22patch-install}

\begin{itemize}
\item
  Follow this link:
  \href{https://support.xilinx.com/s/article/76960?language=en_US}{https ://support.xilinx.com/s/article/76960?language=en\_US}
\item
  Download the file at the bottom of th page and unzip it in
  \texttt{\$XILINX\_ROOT\_DIR}
\item
  run \texttt{cd\ \$XILINX\_ROOT\_DIR}
\item
  run (in one single command line):\\
  \texttt{export\ LD\_LIBRARY\_PATH=\$PWD/Vivado/ $\backslash$} \\
  \texttt{\ \ \ \ \ \ \ \ 2020.2/tps/lnx64/python-3.8.3/lib/ $\backslash$}\\
  \texttt{\ \ \ \ \ \ \ \ Vivado/2020.2/tps/lnx64/python-3.8.3/bin/python3\ y2k22\_patch/patch.py}
\end{itemize}




\section{Use Syfala (clone and launch)}
\label{sec-syfala}
The syfala repository is freely accessible (reading only) on  github (\url{https://github.com/inria-emeraude/syfala}), you have to have a github account of course to clone it. As mentionned before, there may be several sub-directories with different version of Syfala (i.e. different interface for Faust hardware IP). Here are the step needed to run Syfala (after having following the installation instruction of Sections above):
\begin{itemize}
\item Clone the Syfala github repository.
\item install the {\tt syfala.tcl} script
\item Run the script
\end{itemize}

\subsection{Clone the Syfala repository}
to clone the version needed and compile a first architecture you can use the following commands:\\

\begin{boxedminipage}{\textwidth}
  \begin{verbatim}
    git clone https://github.com/inria-emeraude/syfala mysyfala
    cd mysyfala/
    ./syfala.tcl install
    syfala examples/virtualAnalog.dsp
\end{verbatim}
\end{boxedminipage}

~\\

or if you have installed your ssh key on github:\\

\begin{boxedminipage}{\textwidth}
  \begin{verbatim}
    git git@github.com:inria-emeraude/syfala.git mysyfala
    cd mysyfala/
    ./syfala.tcl install
    syfala examples/virtualAnalog.dsp
\end{verbatim}
\end{boxedminipage}


\subsection{Use the {\tt syfala.tcl} script}

the command:

\texttt{\$\ ./syfala.tcl\ install}

will install a
\textbf{symlink} in \textbf{/usr/bin}. After this you'll be able to just
run:

\texttt{\$\ syfala\ myfaustprogram.dsp}

You'll also have to \textbf{edit} your shell \textbf{resource}
\textbf{file} (\textasciitilde/.\textbf{bashrc} /
\textasciitilde/.\textbf{zshrc}) and set the following environment
variable:

\begin{verbatim}
export XILINX_ROOT_DIR=/my/path/to/Xilinx/root/directory
\end{verbatim}

\texttt{XILINX\_ROOT\_DIR} is the root directory where all of the Xilinx
tools (Vivado, Vitis, Vitis\_HLS) are installed.


\subsubsection{Major Syfala commands}\label{quick-start}

\hypertarget{build-examples}{%
\paragraph{build examples}\label{build-examples}}

\begin{lstlisting}
$ syfala examples/virtualAnalog.dsp
# -> runs full toolchain on the virtualAnalog.dsp Faust dsp file, which will be ready to be flashed afterwards on a Zybo Z710 board (by default)

$ syfala examples/virtualAnalog.dsp --board GENESYS --sample-rate 96000
# -> runs full toolchain for the Genesys board, with a sample-rate of 96000Hz

$ syfala examples/phasor.dsp --export phasor-build
# -> runs full toolchain on 'phasor.dsp', automatically exporting the build to 
# export/phasor-build.zip

$ syfala examples/fm.dsp --arch --hls --report
# -> only run 'arch' & 'high-level synthesis' (HLS) step on 'fm.dsp', and show the report afterwards.

$ syfala examples/fm.dsp --board Z20 --arch --hls --export z20-fm-hls-build
# -> only run 'arch' & HLS step on 'fm.dsp' for Zybo Z20 board, and export the build.
\end{lstlisting}

\subsubsection{Additional Syfala `one-shot' commands}

\begin{tabular}{|c|p{9cm}|c|}
  \toprule
  name & description & arguments \\
\midrule
\texttt{install} & installs this script as a symlink in /usr/bin/ &
none \\
\texttt{clean} & deletes current build directory & none \\
\texttt{import} & sets previously exported build as the
current build & .zip target\\
\texttt{export} & exports current build in a .zip file located in the
`export' directory & build name\\
\texttt{report} & prints HLS report of the current build & none \\
\texttt{demo} & fully builds demo based on default example
(virtualAnalog.dsp) & none \\
\texttt{flash} & flashes current build onto target device & none \\
\texttt{gui} & executes the Faust-generated gui application & none \\
\texttt{rebuild-app} & rebuilds the host control application, without
re-synthesizing the whole project & none \\
\texttt{open-project} & opens the generated .xpr project
with Vivado & none \\

\bottomrule
\end{tabular}

\paragraph{Syfala `one-shot' command examples}

\begin{verbatim}
$ syfala clean
$ syfala demo
$ syfala export my-current-build
$ syfala rebuild-app
$ syfala flash
\end{verbatim}

\subsubsection{General Options to Syfala command}

\begin{tabular}{|c|c|p{8cm}|}
  \toprule
option & accepted values & description \\
\midrule
\texttt{-c\ -\/-compiler} & \texttt{HLS\ -\ VHDL} & chooses between
Vitis HLS and faust2vhdl for DSP IP generation. \\
\texttt{-\/-reset} & / & resets current build directory before building
(\textbf{careful}! all files from previous build will be lost) \\
\bottomrule
\end{tabular}

\subsubsection{Controling Syfala Run steps}

\textbf{Note}: the \texttt{-\/-all} is not necessary if you wish to run
all steps, just run:

\texttt{syfala\ myfaustdsp.dsp}

\begin{tabular}{|c|p{12cm}|}
  \toprule
\texttt{-\/-all} & runs all toolchain compilation steps (from
\texttt{-\/-arch} to \texttt{-\/-gui}) \\
\midrule
\texttt{-\/-arch} & uses Faust to generate ip/host cpp files for HLS and
Host application compilation \\
\texttt{-\/-hls\ -\/-ip} & runs Vitis HLS on generated ip cpp file \\
\texttt{-\/-project} & generates Vivado project \\
\texttt{-\/-synth} & synthesizes full Vivado project \\
\texttt{-\/-host\ -\/-app} & compiles Host application, exports sources
and .elf output to \texttt{build/sw\_export} \\
\texttt{-\/-gui} & compiles Faust GUI controller \\
\texttt{-\/-flash} & flashes boot files on device at the end of the
run \\
\texttt{-\/-report} & prints HLS report at the end of the run \\
\texttt{-\/-export} & \texttt{\textless{}id\textgreater{}} exports build
to export/ directory at the end of the run \\
\bottomrule
\end{tabular}

\subsubsection{Controlling the architecture build by Syfala}

\begin{tabular}{|c|c|c|}
  \toprule
parameter & accepted values & default value \\
\midrule
\texttt{-\/-memory,\ -m} & \texttt{DDR\ -\ STATIC} & \texttt{DDR} \\
\texttt{-\/-board,\ -b} & \texttt{Z10\ -\ Z20\ -\ GENESYS} &
\texttt{Z10} \\
\texttt{-\/-sample-rate} &
\texttt{48000\ -\ 96000\ -\ 192000\ -\ 384000\ -\ 768000} &
\texttt{48000} \\
\texttt{-\/-sample-width} & \texttt{16\ -\ 24\ -\ 32} & \texttt{24} \\
\texttt{-\/-controller-type} &
\texttt{DEMO\ -\ PCB1\ -\ PCB2\ -\ PCB3\ -\ PCB4} & \texttt{PCB1} \\
\texttt{-\/-ssm-volume} & \texttt{FULL\ -\ HEADPHONE\ -\ DEFAULT} &
\texttt{DEFAULT} \\
\texttt{-\/-ssm-speed} & \texttt{FAST\ -\ DEFAULT} & \texttt{DEFAULT} \\
\bottomrule
\end{tabular}
\\

Here is the description of these parameters:\\
\begin{tabular}{|c|p{12cm}|}
  \toprule
parameter & description \\
\midrule
\texttt{-\/-memory,\ -m} & selects if \textbf{external} \textbf{DDR3} is
used. Enable if you use some delay, disable if you do want any memory
access (should not be disabled) \\
\texttt{-\/-board} & Defines target board. \textbf{Z10} ,\textbf{Z20}
and \textbf{GENESYS} only. If you have a VGA port (rather than 2 HDMI
ports), you have an old Zybo version, which is not supported. \\
\texttt{-\/-sample-rate} & Changes \textbf{sample rate} value (Hz). Only
48kHz and 96kHz is available for \textbf{SSM} embeded codec. 192000
(\textbf{ADAU1777} and \textbf{ADAU1787} only) 384000 (\textbf{ADAU1787}
only) 768000 (\textbf{ADAU1787} only and with
\texttt{-\/-sample-\/-width\ 16} only) \\
\texttt{-\/-sample-width} & Defines \textbf{sample bit depth}
(16\textbar24\textbar32) \\
\texttt{-\/-controller-type} & Defines the controller used to drive the
controls when \textbf{SW3} is \textbf{UP}. (\textbf{SW3} \textbf{DOWN}
for \textbf{software} control), \textbf{SEE BELOW} for details on each
value \\
\texttt{-\/-ssm-volume} & Chooses audio codec to use. For now, it only
changes the scale factor. \textbf{FULL}: Maximum (\textbf{WARNING}: for
speaker only, do not use with headphones). \textbf{HEADPHONE}: Lower
volume for headphone use. \textbf{DEFAULT}: Default value +1dB because
the true 0dB (\texttt{0b001111001}) decreases the signal a little
bit. \\
\texttt{-\/-ssm-speed} & Changes \textbf{SSM ADC/DAC} sample rate.
\textbf{DEFAULT}: 48kHz sample rate. \textbf{FAST}: 96Khz sample rate \\
\bottomrule
\end{tabular}

\section{Hardware configuration (Zybo Z7-10/20)}
\label{hard}
\begin{itemize}

\item
  Jumper \textbf{JP5} should be on \emph{JTAG}
\item
  \textbf{Power select} jumper should be on \emph{USB}\\
\item
  \textbf{Switches} SW0, SW1, SW2, SW3 should be \textbf{down} (i.e. toward the opposite side of the ethernet connector\\
\item
  The \textbf{audio input} is \textbf{LINE IN} (blue), not MIC IN\\
\item
  The \textbf{audio output} is the black \textbf{HPH OUT} jack
\end{itemize}

\subsection{Control of the Syfala IP}
\label{control}

To control your DSP, you can either use a Syfala Hardware Controller Board or a
GUI on your computer. Beguinner should use GUI control.

\hypertarget{gui-sw3-down}{%
\paragraph{GUI (SW3 DOWN)}\label{gui-sw3-down}}

\textbf{SW3} should be \textbf{down} (0).

If you use GUI, open the GUIcontroller after booting with the following
command:

\begin{verbatim}
make gui
\end{verbatim}

\hypertarget{syfala-hardware-controller-board-sw3-up}{%
\paragraph{Syfala Hardware Controller Board (SW3
UP)}\label{syfala-hardware-controller-board-sw3-up}}

\textbf{SW3} should be \textbf{up} .

If you use a Hardware Controller Board, please set the
\texttt{-\/-controller-type} command-line parameter to the proper value
(see below)

\hypertarget{controller-type-values-description}{%
\subparagraph{Controller-type values
description}\label{controller-type-values-description}}

\begin{itemize}

\item
  \textbf{DEMO}: Popophone demo box
\item
  \textbf{PCB1}: Emeraude PCB config 1: 4 knobs, 2 switches, 2 sliders
  (default)
\item
  \textbf{PCB2}: Emeraude PCB config 2: 8 knobs
\item
  \textbf{PCB3}: Emeraude PCB config 3: 4 knobs, 4 switches
\item
  \textbf{PCB4}: Emeraude PCB config 4: 4 knobs above, 4 switches below
\end{itemize}

You can swap from hardware to software controller during DSP execution
by changing SW3.

\hypertarget{switch-description}{%
\subsubsection{Switch description}\label{switch-description}}

\begin{verbatim}
  SW3   SW2    SW1    SW0
+-----+-----+-------+------+
| Hard| ADAU| BYPASS| MUTE |
|     |     |       |      |
|     |     |       |      |
| GUI | SSM |USE DSP|UNMUTE|
+-----+-----+-------+------+
\end{verbatim}
\begin{itemize}

\item
  \textbf{SW3}: Controller type select: hardware (Controller board) or
  software (GUI). Default: {\bf GUI}
\item
  \textbf{SW2}: Audio codec input select (ADAU=external or SSM=onboard).
  Does not affect output. Default: \textbf{SSM}
\item
  \textbf{SW1}: Bypass audio dsp. Default: \textbf{USE DSP}
\item
  \textbf{SW0}: Mute. Default: \textbf{UNMUTE}
\end{itemize}

\hypertarget{status-leds}{%
\subsubsection{Status LEDs}\label{status-leds}}

The RGB led indicate the program state:

\begin{itemize}

\item
  \textbf{BLUE} = WAITING
\item
  \textbf{GREEN} = ALL GOOD
\item
  \textbf{ORANGE} = WARNING (Bypass or mute enable)
\item
  \textbf{RED} = ERROR (Config failed or incompatible). Could happen if
  you select SSM codec with incompatible sample rate.
\end{itemize}

The 4 LEDs above the switches indicate the switches state. If one of
them blink, it indicates the source of the warning/error.

\hypertarget{sd-card-files}{%
\subsubsection{SD card files}\label{sd-card-files}}

You can put the program on an SD card (if you want something
reproductible and easily launchable, for the demos\ldots).\\
After a \texttt{make} command, you should see a \texttt{BOOT.bin} file
in SW\_export (or you can build it with \texttt{make\ boot\_file}).\\
Put the file on the root of SD card. And don't forget to put JP5 on `SD'
position !




\input{known-bugs}

\input{syfala-team}
